---
title: Validation of ICD-9-CM/ICD-10-CM Codes for Automated Electronic Scoring of
  APACHE II, APACHE III, and SAPS II
subtitle: Analysis for ASHP Poster
author: Eric Kao, Brian Gulbis
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(forcats)
library(tableone)
library(broom)

dirr::get_rds("../data/final")
```

## Patient Selection

* Patients meeting inclusion criteria: `r exclude$screen`
* Reasons for exclusion:
    - Pregnant: `r exclude$pregnant`
    - Prisoner: `r exclude$prisoners`
    - ICU stay < 12 hours: `r exclude$icu_short`
    - Diagnosis coding contained both ICD-9-CM and ICD-10-CM: `r exclude$mult_icd_types`
    - Missing data needed to calculate risk scores: `r exclude$labs_missing`
    - Re-encounters (only included each patient once) : `r exclude$reencounter`
* Total number of eligible patients remaining to sample from: `r flatten_int(exclude)[1] - sum(flatten_int(exclude)[-1])`

## Patient Demographics

```{r}
deceased <- "Deceased/Hospice"
home <- "Home"
transfer <- "Transferred"

dispos <- list("Deceased" = deceased,
               "Expired/Donor" = deceased, 
               "Hospice-Home" = deceased, 
               "Hospice-Medical Facility" = deceased,
               "Home Care with Home Health" = home,
               "Home or Self Care" = home,
               "Left Against Medical Advise" = home,
               "Skilled Nursing Facility" = transfer,
               "DC/DISC TO REHAB" = transfer,
               "Long Term Care" = transfer,
               "Intermediate Care" = transfer,
               "Acute Care" = transfer,
               "Federal Healthcare Facility" = transfer,
               "DC/TF To Psych Hosp" = transfer)

df <- data_demographics %>%
    dmap_at("disposition", str_replace_all, pattern = dispos) %>%
    dmap_at("race", factor) %>%
    dmap_at("race", fct_lump, n = 4) %>%
    select(-pie.id, Age = age, Sex = sex, Race = race, Disposition = disposition, `Hospital Length of Stay (days)` = length.stay, `ICU Location` = location) 

vars <- names(df)
tbl <- CreateTableOne(vars, data = df)
ptbl <- print(tbl, printToggle = FALSE)
rownames(ptbl) <- str_replace_all(rownames(ptbl), "   ", "- ")
knitr::kable(ptbl, caption = "Demographics")
```


## Primary endpoint

### APACHE II

* Mean and median score for each comorbidity set

```{r}
scores <- ls(pattern = "score_apache2")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, apache2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_apache2_", replacement = "") 

man <- filter(all_scores, score == "manual")$apache2

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

```{r}
ggplot(all_scores, aes(x = score, y = apache2)) +
    geom_boxplot()
```


* Comparison of scores
    - Results run using both the t-test and Wilcoxon signed rank test (assumes non-parametric data, is more difficult to achieve significance)
    - To account for performing multiple comparisons, a BH (Benjamini and Hochberg) correction is applied to the p-values; **use the bh.adjust value as the p-value**
        + Reference: http://ies.ed.gov/ncee/pubs/20084018/app_b.asp

```{r}
t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$apache2, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(t, digits = 4)
```

```{r}
mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$apache2, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(mw, digits = 4)
```

* Compare differences in score from comorbidity sets vs. manual
    - Column *num_diff* indicates the number of patients with a score that differed from manual
    - Mean and median indicate the difference from the manual score
        + These were only calculated for those patients whose score was different from manual
        + If all patients were included, the number is skewed towards 0 by all the patients with the same scores

```{r}
diff <- all_scores %>%
    spread(score, apache2) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, apache2, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(apache2 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Differences in APACHE II Scores")
```

### APACHE III

```{r}
scores <- ls(pattern = "score_apache3")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, apache3)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_apache3_", replacement = "") 

man <- filter(all_scores, score == "manual")$apache3

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

```{r}
ggplot(all_scores, aes(x = score, y = apache3)) +
    geom_boxplot()
```

* Comparison of scores

```{r}
t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$apache3, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(t, digits = 4)
```

```{r}
mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$apache3, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(mw, digits = 4)
```

* Differences in score

```{r}
diff <- all_scores %>%
    spread(score, apache3) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, apache3, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(apache3 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Difference in APACHE III Scores")
```

### SAPS II

```{r}
scores <- ls(pattern = "score_saps2")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, saps2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_saps2_", replacement = "") 

man <- filter(all_scores, score == "manual")$saps2

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

```{r}
ggplot(all_scores, aes(x = score, y = saps2)) +
    geom_boxplot()
```

* Comparison of scores

```{r}
t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$saps2, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(t, digits = 4)
```

* Confidence interval was unable to be calculated for Wilcoxon analysis here

```{r}
mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$saps2, man, paired = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(mw, digits = 4)
```

* Differences in score

```{r}
diff <- all_scores %>%
    spread(score, saps2) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, saps2, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(saps2 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Difference in SAPS II Scores")
```

## Secondary endpoints

### Predicted mortality

#### SAPS II

* Mean and median risk of mortality (as a percent) for each comorbidity set

```{r}
comorbid <- ls(pattern = "score_saps2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, saps2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_saps2_", replacement = "") %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    mutate(logit = -7.763 + 0.0737 * saps2 + 0.9971 * log(saps2 + 1),
           mortality = exp(logit) / (1 + exp(logit))) %>%
    select(pie.id, score, mortality) 

man <- filter(all_df, score == "manual")$mortality

result <- all_df %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result, digits = 4)
```

```{r}
t <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>% 
    do(tidy(t.test(.$mortality, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(t, digits = 4)
```

```{r}
mw <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$mortality, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(mw, digits = 4)
```

* Difference in risk of mortality
    - Mean and median calculated using only those patients whose predicted mortality score was different from manual
    
```{r}
diff <- all_df %>%
    spread(score, mortality) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, saps2, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(saps2 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Difference in Predicted Mortality using SAPS II")
```

### Number of comorbidities

#### APACHE II

* Mean represents the average number of comorbidities per patient

```{r}
comorbid <- ls(pattern = "data_apache2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, liver, cardiovasc, respiratory, renal, immunocomp, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache2_", replacement = "") %>%
    mutate_if(is.logical, as.numeric) %>%
    by_row(function(x) sum(x[, 3:7], na.rm = TRUE), .collate = "rows", .to = "num_comorbid") %>%
    select(pie.id, score, num_comorbid) 

man <- filter(all_df, score == "manual")$num_comorbid

result <- all_df %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

```{r}
t <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$num_comorbid, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(t, digits = 4)
```

```{r}
mw <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$num_comorbid, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.adjust = p.adjust(p.value, "BH")) %>%
    select(score:p.value, bh.adjust, everything())

knitr::kable(mw, digits = 4)
```

### Proportion with differing comorbidity

#### APACHE II

* Proportion of patients where a comorbidity was identified by ICD or by manual, but not both

```{r}
comorbid <- ls(pattern = "data_apache2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache2_", replacement = "") %>%
    spread(score, comorbidity) %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    ungroup() %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean))

knitr::kable(all_df)
```

#### APACHE III

* Proportion of patients where the primary comorbidity identified by ICD varied from primary comorbidity identified by manual

```{r}
comorbid <- ls(pattern = "data_apache3_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache3_", replacement = "") %>%
    spread(score, comorbidity, fill = "") %>%
    dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean))

knitr::kable(all_df)
```

#### SAPS II

* Proportion of patients where the primary comorbidity identified by ICD varied from primary comorbidity identified by manual

```{r}
comorbid <- ls(pattern = "data_saps2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_saps2_", replacement = "") %>%
    spread(score, comorbidity, fill = "") %>%
    dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean))

knitr::kable(all_df)
```
