---
title: Validation of ICD-9-CM/ICD-10-CM Codes for Automated Electronic Scoring of
  APACHE II, APACHE III, and SAPS II
subtitle: Preliminary Analysis for ASHP Abstract Submission
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
# library(tableone)
library(broom)

dirr::get_rds("../data/final")
```

## Patient Selection

* Patients meeting inclusion criteria: `r exclude$screen`
* Reasons for exclusion:
    - Pregnant: `r exclude$pregnant`
    - Prisoner: `r exclude$prisoners`
    - ICU stay < 12 hours: `r exclude$icu_short`
    - Diagnosis coding contained both ICD-9-CM and ICD-10-CM: `r exclude$mult_icd_types`
    - Missing data needed to calculate risk scores: `r exclude$labs_missing`
    - Re-encounters (only included each patient once) : `r exclude$reencounter`
* Total number of eligible patients remaining to sample from: `r flatten_int(exclude)[1] - sum(flatten_int(exclude)[-1])`

## Outcomes

### Primary endpoint

#### APACHE II

* Mean and median score for each comorbidity set

```{r}
scores <- ls(pattern = "score_apache2")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, apache2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "score_apache2_", replacement = "") %>%
    arrange(pie.id, score)

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

* Comparison of scores
    - Results run using both the t-test and Wilcoxon signed rank test (assumes non-parametric data, is more difficult to achieve significance)

```{r}
man <- filter(all_scores, score == "manual")$apache2

t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$apache2, man, paired = TRUE)))

knitr::kable(t)
```

```{r}
mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$apache2, man, paired = TRUE, conf.int = TRUE, exact = FALSE)))

knitr::kable(mw)
```

* Compare differences in score from comorbidity sets vs. manual
    - Column *num_diff* indicates the number of patients with a score that differed from manual
    - Mean and median indicate the difference from the manual score
        + These were only calculated for those patients whose score was different from manual
        + If all patients were included, the number is skewed towards 0 by all the patients with the same scores

```{r}
diff <- all_scores %>%
    spread(score, apache2) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, apache2, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(apache2 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Differences in APACHE II Scores")
```

#### APACHE III

```{r}
scores <- ls(pattern = "score_apache3")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, apache3)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "score_apache3_", replacement = "") %>%
    arrange(pie.id, score)

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

* Comparison of scores

```{r}
man <- filter(all_scores, score == "manual")$apache3

t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$apache3, man, paired = TRUE)))

knitr::kable(t)
```

```{r}
mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$apache3, man, paired = TRUE, conf.int = TRUE, exact = FALSE)))

knitr::kable(mw)
```

* Differences in score

```{r}
diff <- all_scores %>%
    spread(score, apache3) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, apache3, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(apache3 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Difference in APACHE III Scores")
```

#### SAPS II

```{r}
scores <- ls(pattern = "score_saps2")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, saps2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "score_saps2_", replacement = "") %>%
    arrange(pie.id, score)

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

* Comparison of scores

```{r}
man <- filter(all_scores, score == "manual")$saps2

t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$saps2, man, paired = TRUE)))

knitr::kable(t)
```

* Confidence interval was unable to be calculated for Wilcoxon analysis here

```{r}
mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$saps2, man, paired = TRUE, exact = FALSE)))

knitr::kable(mw)
```

* Differences in score

```{r}
diff <- all_scores %>%
    spread(score, saps2) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, saps2, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(saps2 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Difference in SAPS II Scores")
```

### Secondary endpoints

#### Difference in predicted mortality using SAPS II

* Mean and median risk of mortality (as a percent) for each comorbidity set

```{r}
comorbid <- ls(pattern = "score_saps2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, saps2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "score_saps2_", replacement = "") %>%
    arrange(pie.id, score) %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    mutate(logit = -7.763 + 0.0737 * saps2 + 0.9971 * log(saps2 + 1),
           mortality = exp(logit) / (1 + exp(logit))) %>%
    select(pie.id, score, mortality) 

result <- all_df %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

```{r}
man <- filter(all_df, score == "manual")$mortality

t <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$mortality, man, paired = TRUE)))

knitr::kable(t)
```

```{r}
mw <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$mortality, man, paired = TRUE, conf.int = TRUE, exact = FALSE)))

knitr::kable(mw)
```

* Difference in risk of mortality
    - Mean and median calculated using only those patients whose predicted mortality score was different from manual
    
```{r}
diff <- all_df %>%
    spread(score, mortality) %>%
    mutate_at(vars(-pie.id, -manual), funs(abs(. - manual))) %>%
    gather(score, saps2, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    filter(saps2 > 0) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = length, mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff, caption = "Difference in Predicted Mortality using SAPS II")
```

#### Difference in number of comorbidities identified using APACHE II

```{r}
comorbid <- ls(pattern = "data_apache2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, liver:comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache2_", replacement = "") %>%
    arrange(pie.id, score) %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    mutate_if(is.logical, as.numeric) %>%
    by_row(function(x) sum(x[, 3:7], na.rm = TRUE), .collate = "rows", .to = "num_comorbid") %>%
    select(pie.id, score, num_comorbid) 

result <- all_df %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(result)
```

```{r}
man <- filter(all_df, score == "manual")$num_comorbid

t <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$num_comorbid, man, paired = TRUE)))

knitr::kable(t)
```

* Could not calculate confidence intervals

```{r}
mw <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$num_comorbid, man, paired = TRUE, exact = FALSE)))

knitr::kable(mw)
```

```{r}
diff <- all_df %>%
    select(-pie.id) %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    dmap_at("num_comorbid", ~ abs(.x - man)) %>%
    filter(num_comorbid > 0) %>%
    summarize_all(funs(num_diff = sum(. > 0), mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75)))

knitr::kable(diff)
```

#### Proportion of patients with differing identification of any comorbidity using APACHE II

```{r}
comorbid <- ls(pattern = "data_apache2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache2_", replacement = "") %>%
    arrange(pie.id, score) %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    spread(score, comorbidity, fill = FALSE) %>%
    # dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    ungroup() %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean))

knitr::kable(all_df)
```


#### Proportion of patients with a differing primary comorbidity using APACHE III

```{r}
comorbid <- ls(pattern = "data_apache3_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache3_", replacement = "") %>%
    arrange(pie.id, score) %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    spread(score, comorbidity, fill = "") %>%
    dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    ungroup() %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean))

knitr::kable(all_df)
```

#### Proportion of patients with a differing primary comorbidity using SAPS II

```{r}
comorbid <- ls(pattern = "data_saps2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    dmap_at("score", str_replace_all, pattern = "data_saps2_", replacement = "") %>%
    arrange(pie.id, score) %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    spread(score, comorbidity, fill = "") %>%
    dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    ungroup() %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean))

knitr::kable(all_df)
```
