---
title: "Validation of ICD-9-CM/ICD-10-CM Codes for Automated Electronic Scoring of APACHE II, APACHE III, and SAPS II"
subtitle: "Supplemental Material"
author: "Eric Kao, Brian Gulbis"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  tufte::tufte_html: default
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'), echo = FALSE)
options(htmltools.dir.version = FALSE)

library(tidyverse)
library(stringr)
library(forcats)
library(tableone)
library(broom)
library(icd)
library(icuriskr)

dirr::get_rds("../data/final")
```

# Patient Selection

There were `r format(exclude$screen, big.mark = ",")` patient encounters identified which met inclusion criteria. Patients were then excluded for the following reasons:

* Pregnant: `r format(exclude$pregnant, big.mark = ",")`
* Prisoner: `r format(exclude$prisoners, big.mark = ",")`
* Missing data needed to calculate risk scores: `r format(exclude$labs_missing + exclude$mult_icd_types + exclude$icu_short, big.mark = ",")`
* Re-encounters (each patient  once) : `r format(exclude$reencounter, big.mark = ",")`

This left a total of `r format(flatten_int(exclude)[1] - sum(flatten_int(exclude)[-1]), big.mark = ",")` eligible patients remaining. From these, a random sample of 200 patients were selected for inclusion in the study. 

# Comorbidity Definitions

The following tables contain the list of ICD-9-CM and ICD-10-CM codes which were included in the author-defined comorbidity set.^[Any child-codes for the codes listed were also included (i.e., 196 includes 196.0, 196.1, etc.)] 

```{r, message=FALSE, warning=FALSE}
icd9 <- comorbidity_map_icd9_ek %>%
    map(sort) %>%
    map(icd_condense, warn = FALSE) %>%
    map(icd_short_to_decimal) %>%
    map(str_c, collapse = ", ") %>%
    simplify() %>%
    as_tibble() %>%
    rownames_to_column("comorbidity")

defs <- read_csv("comorbidity_definitions.csv")

df <- defs %>%
    filter(score == "apache2") %>%
    left_join(icd9, by = c("icd" = "comorbidity")) %>%
    select(Group = group, Comorbidity = comorbidity, Criteria = criteria, `ICD-9-CM Code` = value) %>%
    dmap(~ coalesce(.x, ""))

knitr::kable(df, caption = "ICD-9-CM Codes Used in Author-Defined Comorbidity Set for APACHE II")
```

```{r}
icd10 <- comorbidity_map_icd10_ek %>%
    # map(icd_condense, warn = FALSE) %>%
    map(sort) %>%
    map(icd_short_to_decimal) %>%
    map(str_c, collapse = ", ") %>%
    simplify() %>%
    as_tibble() %>%
    rownames_to_column("comorbidity") 

df <- defs %>%
    filter(score == "apache2") %>%
    left_join(icd10, by = c("icd" = "comorbidity")) %>%
    select(Group = group, Comorbidity = comorbidity, Criteria = criteria, `ICD-10-CM Code` = value) %>%
    dmap(~ coalesce(.x, ""))

knitr::kable(df, caption = "ICD-10-CM Codes Used in Author-Defined Comorbidity Set for APACHE II")
```

```{r}
df <- defs %>%
    filter(score == "apache3") %>%
    left_join(icd9, by = c("icd" = "comorbidity")) %>%
    select(Comorbidity = comorbidity, Criteria = criteria, `ICD-9-CM Code` = value) %>%
    dmap(~ coalesce(.x, ""))

knitr::kable(df, caption = "ICD-9-CM Codes Used in Author-Defined Comorbidity Set for APACHE III")
```

```{r}
df <- defs %>%
    filter(score == "apache3") %>%
    left_join(icd10, by = c("icd" = "comorbidity")) %>%
    select(Comorbidity = comorbidity, Criteria = criteria, `ICD-10-CM Code` = value) %>%
    dmap(~ coalesce(.x, ""))

knitr::kable(df, caption = "ICD-10-CM Codes Used in Author-Defined Comorbidity Set for APACHE III")
```

```{r}
df <- defs %>%
    filter(score == "saps2") %>%
    left_join(icd9, by = c("icd" = "comorbidity")) %>%
    select(Comorbidity = comorbidity, Criteria = criteria, `ICD-9-CM Code` = value) %>%
    dmap(~ coalesce(.x, ""))

knitr::kable(df, caption = "ICD-9-CM Codes Used in Author-Defined Comorbidity Set for SAPS II")
```

```{r}
df <- defs %>%
    filter(score == "saps2") %>%
    left_join(icd10, by = c("icd" = "comorbidity")) %>%
    select(Comorbidity = comorbidity, Criteria = criteria, `ICD-10-CM Code` = value) %>%
    dmap(~ coalesce(.x, ""))

knitr::kable(df, caption = "ICD-10-CM Codes Used in Author-Defined Comorbidity Set for SAPS II")
```


# Statistical Methods

Data processing, risk score calculations, and statistical analysis were performed using `r R.Version()$version.string` (@R-base). The `icd` package (@R-icd) was used to cross-reference patient diagnosis codes with comorbidity set diagnosis codes.

Continuous data were analyzed using a paired t-test and Wilcoxon signed rank test. The p-value was then adjusted using the Benjamini and Hochberg method to account for performing multiple comparisons. 

A univarite linear regression model was fit using the risk score from each comorbidity set to predict the risk score for the manual review set. 

# Supplemental Results

The following tables contain the mean and median risk score calculated using each of the eight comorbidity sets, as well as the manual review set. A comparison of the means and medians was performed.^[The adjusted p-value and 95% confidence intervals (CI) are reported for the t-test and Wilcoxon rank sum (denoted by the *)]

## Primary Endpoint

```{r}
scores <- ls(pattern = "score_apache2")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, apache2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_apache2_", replacement = "") 

man <- filter(all_scores, score == "manual")$apache2

score_type <- list("ahrq" = "AHRQ",
                   "ek" = "Authors",
                   "elixhauser" = "Elixhauser",
                   "manual" = "Manual",
                   "quan" = "Quan",
                   "_drg" = "+DRG")

t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$apache2, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.t = p.adjust(p.value, "BH"),
           ci.t = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$apache2, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.mw = p.adjust(p.value, "BH"),
           ci.mw = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75))) %>%
    mutate(IQR = paste(IQR_low, IQR_high, sep = " - ")) %>%
    left_join(t, by = "score") %>%
    left_join(mw, by = "score") %>%
    dmap_at("score", str_replace_all, pattern = score_type) %>%
    select(Tool = score, Mean = mean, SD = sd, `P-value` = bh.t, `95% CI` = ci.t, Median = median, IQR, `P-value*` = bh.mw, `95% CI*` = ci.mw) 
    
knitr::kable(result, digits = 3, align = "lccccccc", caption = "APACHE II Score")
```

```{r}
scores <- ls(pattern = "score_apache3")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, apache3)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_apache3_", replacement = "") 

man <- filter(all_scores, score == "manual")$apache3

t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$apache3, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.t = p.adjust(p.value, "BH"),
           ci.t = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$apache3, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.mw = p.adjust(p.value, "BH"),
           ci.mw = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75))) %>%
    mutate(IQR = paste(round(IQR_low, 1), round(IQR_high, 1), sep = " - ")) %>%
    left_join(t, by = "score") %>%
    left_join(mw, by = "score") %>%
    dmap_at("score", str_replace_all, pattern = score_type) %>%
    select(Tool = score, Mean = mean, SD = sd, `P-value` = bh.t, `95% CI` = ci.t, Median = median, IQR, `P-value*` = bh.mw, `95% CI*` = ci.mw) 


knitr::kable(result, digits = 3, align = "lccccccc", caption = "APACHE III Score")
```

```{r}
scores <- ls(pattern = "score_saps2")

all_scores <- map(scores, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, saps2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_saps2_", replacement = "") 

man <- filter(all_scores, score == "manual")$saps2

t <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$saps2, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.t = p.adjust(p.value, "BH"),
           ci.t = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

mw <- all_scores[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$saps2, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.mw = p.adjust(p.value, "BH"),
           ci.mw = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

result <- all_scores %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75))) %>%
    mutate(IQR = paste(round(IQR_low, 1), round(IQR_high, 1), sep = " - ")) %>%
    left_join(t, by = "score") %>%
    left_join(mw, by = "score") %>%
    dmap_at("score", str_replace_all, pattern = score_type) %>%
    select(Tool = score, Mean = mean, SD = sd, `P-value` = bh.t, `95% CI` = ci.t, Median = median, IQR, `P-value*` = bh.mw, `95% CI*` = ci.mw) 


knitr::kable(result, digits = 3, align = "lccccccc", caption = "SAPS II Score")
```

## Secondary Endpoints

```{r}
comorbid <- ls(pattern = "score_saps2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~select(.x, pie.id, score, saps2)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "score_saps2_", replacement = "") %>%
    distinct(pie.id, score, .keep_all = TRUE) %>%
    mutate(logit = -7.763 + 0.0737 * saps2 + 0.9971 * log(saps2 + 1),
           mortality = exp(logit) / (1 + exp(logit))) %>%
    select(pie.id, score, mortality) 

man <- filter(all_df, score == "manual")$mortality

t <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>% 
    do(tidy(t.test(.$mortality, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.t = p.adjust(p.value, "BH"),
           ci.t = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

mw <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$mortality, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.mw = p.adjust(p.value, "BH"),
           ci.mw = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

result <- all_df %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75))) %>%
    mutate(IQR = paste(round(IQR_low, 1), round(IQR_high, 1), sep = " - ")) %>%
    left_join(t, by = "score") %>%
    left_join(mw, by = "score") %>%
    dmap_at("score", str_replace_all, pattern = score_type) %>%
    select(Tool = score, Mean = mean, SD = sd, `P-value` = bh.t, `95% CI` = ci.t, Median = median, IQR, `P-value*` = bh.mw, `95% CI*` = ci.mw) 

knitr::kable(result, digits = 3, align = "lccccccc", caption = "Predicted Mortality using SAPS II")
```

```{r}
comorbid <- ls(pattern = "data_apache2_")

all_df <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, liver, cardiovasc, respiratory, renal, immunocomp, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache2_", replacement = "") %>%
    mutate_if(is.logical, as.numeric) %>%
    by_row(function(x) sum(x[, 3:7], na.rm = TRUE), .collate = "rows", .to = "num_comorbid") %>%
    select(pie.id, score, num_comorbid) 

man <- filter(all_df, score == "manual")$num_comorbid

t <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(t.test(.$num_comorbid, man, paired = TRUE))) %>%
    ungroup() %>%
    mutate(bh.t = p.adjust(p.value, "BH"),
           ci.t = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

mw <- all_df[, -1] %>%
    filter(score != "manual") %>%
    group_by(score) %>%
    do(tidy(wilcox.test(.$num_comorbid, man, paired = TRUE, conf.int = TRUE, exact = FALSE))) %>%
    ungroup() %>%
    mutate(bh.mw = p.adjust(p.value, "BH"),
           ci.mw = paste(round(conf.low, 3), round(conf.high, 3), sep = " - ")) 

result <- all_df %>%
    select(-pie.id) %>%
    group_by(score) %>%
    summarize_all(funs(mean, sd, median, IQR_low = quantile(., 0.25), IQR_high = quantile(., 0.75))) %>%
    mutate(IQR = paste(round(IQR_low, 1), round(IQR_high, 1), sep = " - ")) %>%
    left_join(t, by = "score") %>%
    left_join(mw, by = "score") %>%
    dmap_at("score", str_replace_all, pattern = score_type) %>%
    select(Tool = score, Mean = mean, SD = sd, `P-value` = bh.t, `95% CI` = ci.t, Median = median, IQR, `P-value*` = bh.mw, `95% CI*` = ci.mw) 

knitr::kable(result, digits = 3, align = "lccccccc", caption = "Number of Comorbidities per Patient using APACHE II")
```

The following table contains the number of patients with different comorbidities identified by the comorbidity sets when compared with manual review. For APACHE II, this was counted as the number of patients where a comorbidity was identified by ICD or by manual, but not both. For APACHE III and SAPS II, this was counted as the number of patients where the primary comorbidity identified by ICD coding varied from the primary comorbidity identified by manual review.

```{r}
comorbid <- ls(pattern = "data_apache2_")

apache2_diff <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache2_", replacement = "") %>%
    spread(score, comorbidity) %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    ungroup() %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean)) %>%
    mutate(apache2_n = paste0(num_diff, " (", round(proportion * 100, 1), ")"))

comorbid <- ls(pattern = "data_apache3_")

apache3_diff <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_apache3_", replacement = "") %>%
    spread(score, comorbidity, fill = "") %>%
    dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean)) %>%
    mutate(apache3_n = paste0(num_diff, " (", round(proportion * 100, 1), ")"))

comorbid <- ls(pattern = "data_saps2_")

saps2_diff <- map(comorbid, ~mutate(get(.x), score = .x)) %>%
    map(~distinct(.x, pie.id, score, comorbidity)) %>%
    map_df(~semi_join(.x, manual_patients, by = "pie.id")) %>%
    arrange(pie.id, score) %>%
    dmap_at("score", str_replace_all, pattern = "data_saps2_", replacement = "") %>%
    spread(score, comorbidity, fill = "") %>%
    dmap(str_replace_all, pattern = "none", replacement = "") %>%
    mutate_at(vars(-pie.id, -manual), funs(. != manual)) %>%
    mutate_if(is.logical, as.numeric) %>%
    gather(score, comorbidity, -pie.id, -manual) %>%
    select(-pie.id, -manual) %>%
    group_by(score) %>%
    summarize_all(funs(num_diff = sum(. > 0), proportion = mean)) %>%
    mutate(saps2_n = paste0(num_diff, " (", round(proportion * 100, 1), ")"))

df <- apache2_diff %>%
    left_join(apache3_diff, by = "score") %>%
    left_join(saps2_diff, by = "score") %>%
    dmap_at("score", str_replace_all, pattern = score_type) %>%
    select(Tool = score, `APACHE II, n (%)` = apache2_n, `APACHE III, n (%)` = apache3_n, `SAPS II, n (%)` = saps2_n)

knitr::kable(df, digits = 3, align = "lccc", caption = "Proportion of Patients with Differing Comorbidities")
```

